The \project{unRAVE} catalog

### 2016-08-24 article/unrave.tex
Text edits and new text on the k2/epic ms training set

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\apogee}{\project{\acronymn{APOGEE}}}
\newcommand{\lamost}{\project{\acronymn{LAMOST}}}
\newcommand{\hipparcos}{\project{Hipparcos}}
\newcommand{\epic}{\project{K2/EPIC}}

@@

\begin{abstract}
The Milky Way may be our greatest laboratory for understanding galaxy formation.
This premise relies on the ability to measure --- for large numbers of stars ---
their detailed chemical abundances, and their 3D positions and motions.  On 14 September
2016 the \gaia\ mission will release the latter for $\approx2\times10^6$ stars.
However, a catalog of precise chemical abundances to complement that data
release is currently lacking.  The RAdial Velocity Experiment (\rave) survey has 
acquired spectra for 292,036 stars in common with the first \gaia\ data release, 
constituting the largest overlap of any stellar spectroscopic sample.  Here we 
perform an independent analysis of \rave\ spectra using an implementation of 
\thecannon\ that uses prior probabilities on atomic line formation and stellar
astrophysical parameters.  Our approach yields precise stellar labels at lower
S/N ratios than most \emph{ab initio} approaches, allowing us to deliver
improved radial velocities, effective temperature $\teff$, surface gravity $\logg$,
and chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) 
for \Nstars\ stars.  We validate our results internally from repeat visits, 
and externally against existing high-resolution spectroscopic surveys.  The 
typical precision in chemical abundances is X.XX~dex.  We provide 
spectrophotometric distances based on our labels, improving upon the parallax
distances available for the fainter targets in the first \gaia\ data release.
\stub{A one-line science result based on RAVE/Hipparcos overlap}
\end{abstract}

@@

The Milky Way is considered to be the best laboratory for understanding galaxy
formation and evolution.  This premise hinges on the ability to precisely measure 
the astrometry and chemistry for (many) individual stars, and to use those data 
to infer the structure, kinematics, and chemical enrichment of the Galaxy 
\citep[e.g.,][]{Schlaufman_2009,Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,
Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Bovy_2016}.  However, these 
quantities are not known for even 1\% of stars in the Milky Way.  Stellar 
distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,Jofre_2015,
Madler_2016}, proper motions can be plagued by unquantified systematics from 
the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
stellar spectroscopists frequently report significantly different chemical 
abundance patterns \citep{Smiljanic_2014}.  The impact this lack of data has on
scientific inferences cannot be understated.  It limits our understanding on a
number of sub-fields in astrophysics, including the properties of exoplanet host
stars, the formation (and destruction) of stars and clusters, as well as the 
structure of the Galactic disk, to name a few.

These formed just some of the scientific drivers for the \gaia\ space telescope.
\gaia\ is primarily an astrometric mission, providing precise positions,
parallaxes and proper motions for more than $10^9$ stars in its final data
release, scheduled for 2022.  While this is a sample size about three orders 
of magnitude larger than its predecessor \hipparcos, both astrometry and 
chemistry are required to fully characterize the formation and evolution of our Galaxy.  
For this reason \gaia\ will also provide radial velocities, stellar parameters 
and chemical abundances for a subset of brighter stars, primarily from the on-board
radial velocity spectrometer.  However, those chemical abundances will not be 
available until later releases.  Until those abundances are available,
astronomers seeking to use chemistry and kinematic measurements are reliant on 
ground-based spectroscopic surveys in order to complement the available astrometry
from \gaia.

The first \project{Gaia} data release will include positions, proper motions, and 
parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} 
catalog \citep{Michalik_2015a,Michalik_2015b}.  After cross-matching all major 
stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2
catalog against the \project{APOGEE} \citep{Zasowski_2013}, \project{Gaia-ESO} 
\citep{Gilmore_2012,Randich_2013}, \project{GALAH} \citep{DeSilva_2015},
\project{LAMOST} \citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006} 
surveys}, we found the RAdial Velocity Experiment (hereafter \rave) survey to 
have the largest overlap with the first \project{Gaia} data release: 292,036 
stars.  We then used the \project{Gaia} universe model \citep{Robin_2012} to 
estimate the precision in parallax and proper motions that will be available in 
\project{Gaia} DR1 for every star in these cross-matches.  Comparing the expected
precision to what is currently available, we further found that the \project{RAVE}
survey will benefit most from \project{Gaia} DR1.  The distances of 63\% of the 
\project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars) are expected to
improve with the first \project{Gaia} data release, and 47\% of stars likely to
have better proper motions (137,211 stars).  This motivated us to examine what 
chemical abundance information was available for \rave, and to evaluate whether 
we could enable new chemodynamic studies by contributing to this existing set of
chemical abundances.

We briefly describe the \rave\ data in section \ref{sec:data}, before explaining
the model we employ in section \ref{sec:model}.  In section \ref{sec:validation}
we outline a number of validation experiments, including: internal sanity checks,
comparisons with literature samples, and investigations to ensure our results
are consistent with expectations from astrophysics.  We provide concluding
remarks in section \ref{sec:conclusion}, as well as a description of how to
access our results electronically.

@@

\rave\ is a magnitude-limited stellar spectroscopic survey of the (nearby) Milky Way,
principally designed to measure radial velocities for up to $10^6$ stars.
Observations were conducted on the 1.2~m UK Schmidt telescope at the Australian 
Astronomical Observatory\footnote{Formerly the Anglo-Australian Observatory} from 
2003--2013.  The large 5.7~degree field-of-view and robotic fibre positioner (6dF)
made for very efficient observing: spectra for up to 150 targets could be 
simultaneously acquired.  When observations concluded in April 2013, a total of 
574,630 spectra had been collected of 483,330 unique objects.

The target selection for \rave\ is based on the $I$-band apparent magnitude,
$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge \citep{Wonjo_2016}.  
The $I$ band was used because it overlaps with the wavelength range that \rave\ 
operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 
near-infrared triplet.  These stellar absorption lines are dominated by pressure 
broadening, which makes them strong and therefore visible even in very low S/N data
or metal-poor stars.  This spectral region is also well-studied.  It is one of the
key setups used for the ground-based high-resolution \project{Gaia-ESO} survey
\citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
coverage is comparable to the Radial Velocity Spectrometer on board the \project{Gaia}
space telescope \citep{Recio-Blanco_2016}.

The exposure times for \rave\ observations were optimised to obtain radial 
velocities for as many stars as possible.  Detailed chemical abundances were
always an important science goal of the survey, but this was a secondary objective.  
For this reason the distribution of S/N ratios in \rave\ spectra is considerably 
lower than other stellar spectroscopic surveys where chemical abundances are the 
primary motivation.  The \rave\ spectra have an effective resolution 
$\mathcal{R} \approx 7500$ and the distribution of S/N ratios peaks at 
$\approx$50~per~pixel$^{-1}$.  For comparison, the \project{GALAH} survey 
\citep{DeSilva_2015} --- which was specifically constructed for detailed chemical 
abundance analyses --- includes a wavelength range about 2.5 times larger at 
resolution $\mathcal{R} \approx 28,000$, and yet the \project{GALAH} project still 
targets for S/N $\gtrsim100$.

@@

Given the high-quality of the normalization performed by the \rave\ team, we chose
not to re-normalize the spectra.  Our tests demonstrated that the procedure 
outlined in \citet{Kunder_2016} is sufficient for our analysis procedure. Therefore,
there is a limited number of pre-processing steps that we conducted before our
analysis could begin.  Firstly, we calculated inverse variance arrays from the
fractional uncertainties provided, then re-sampled the flux and inverse variance
arrays onto a common rest-wavelength map for all stars.  Depending on the fibre 
used and the stellar radial velocity, the range of rest-frame wavelength values
varied for each star.  Given that fluxes were unavailable in the edge pixels for 
most stars, we excluded pixels outside of the rest wavelength range 
$8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.  This corresponds to about
30~pixels excluded on either side of the common wavelength array, leaving us with
945~pixels for science.

@@

Throughout this work we adopt the assumptions and nomenclature first set out in
\citet{Ness_2015}, and extended in \citet{Casey_2016b}.  However, one distinction 
here with \citet{Casey_2016b} is that in this work the \emph{labelled} set and the
\emph{training} set will be the same set.  In \citet{Casey_2016b} two subsets of
the labelled set were defined: the training set and a distinct validation set.
This allowing us to train on the \emph{training set} and then test on the 
\emph{validation set}.  However in this work the labelled set is relatively 
small, about an order of magnitude smaller than what was available in 
\citet{Casey_2016b}, and therefore we will use all labelled set stars in the 
training step and supplement our validation experiments with literature comparisons.
Thus, throughout this work we will use the terms \emph{labelled set} and 
\emph{training set} interchangeably.  Below we describe how we constructed the 
training set, before we outline the model that we adopt.

@@

We sought to construct a training set consisting of stars across the main-sequence,
sub-giant, and the red giant branch.  We require stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \project{Gaia-ESO} survey includes 
giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars incorrectly sit above the isochrones in a classical Hertzsprung-Russell diagram)
If we consider lower-resolution studies, there are 2,369 stars that overlap with 
\project{LAMOST} --- of which 2,213 have positive S/N ratios in the $g$-band 
(\texttt{snrg}).  However, similar systematic effects are seen on the main-sequence: 
the labels are expectedly less precise given the lower resolution, no detailed 
abundance information is available\footnote{Abundance information is available for 
\project{LAMOST} stars from \citet{Ho_2016}, but that sample contains only giant stars},
and there is only sparse sampling across the Hertzsprung-Russell diagram.  The 
\project{GALAH} survey has a good overlap with \rave, but results from that sample 
are not yet precise enough for our purposes.

These constraints forced us to construct a heterogeneous training set.  Given previous
successes in transferring high S/N ratio labels from \project{APOGEE} \citep{Ness_2015,
Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \project{APOGEE}
---\rave\ overlap sample for giant star labels in the training set.  Of these, about
900 are giants according to \project{APOGEE}.  From this sample we selected stars to
have: metallicity determinations ([Fe/H] $> -5$); high S/N ratios in both the 
\project{APOGEE} and \rave\ spectra ($>200$ in \project{APOGEE}; $>50$ in \rave); and
we further required that the \project{ASPCAP} pipeline did not report any peculiar
flags (\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 357 stars along the
giant branch, with metallicities ranging from $[{\rm Fe/H}] = -1.62$ to 0.32.  Our tests
showed that the metallicity range of the training set needed to extend below
$[{\rm Fe/H}] \lesssim -2$, and thus below we describe how we supplemented this
sample with giant stars from the \epic\ catalog \citep{Huber_2016}.

Assembling a training set for the main-sequence and sub-giant branch was more complex.
There are no spectroscopic studies that have a large enough overlap with \rave, and
also cover our required range of stellar parameters (e.g., FGKM-type stars).  Moreover,
most of the spectroscopic studies we considered exhibit a flat lower main-sequence, a
consequence of violating the assumption of local thermal equilibrium 
\citep[e.g., see][for a discussion]{Bensby_2014}.  For these reasons we chose to make
use of the \epic\ catalog \citep{Huber_2016} for the training set labels.  The \epic\
catalog follows from the successful \project{Kepler} input catalog \citep{Brown_2011},
and provides probabilistic stellar classifications for $\gtrsim180{,}000$ stars in 
the \project{K2} fields based on the astrometric, photometric, and spectroscopic
information available.  There are 4,611 stars that overlap between \epic\ and \rave.

When the \epic\ posteriors are narrow (i.e., the uncertainties on labels are small),
\citet{Huber_2016} show that the labels are accurate and trustworthy.  However,
\epic\ differs from the \project{Kepler} input catalog because \epic\ does not
benefit from having narrow-band $DDO_{51}$ photometry in order to aid dwarf/giant
classification.  For this reason there are a number of stars that are mis-classified
in the \epic\ catalog.  \citet{Huber_2016} acknowledges this, noting: the fraction 
of X classified as Y could be as much as Z\%; up to Z2\% of X2 could be marked as Y2, 
and \stub{other things}.  In other words, the \epic\ labels are usually of high 
fidelity, and mis-classifications are clearly detectable if spectra are available.
As we will show, when the mis-classifications can be identified from \rave\ spectra, 
the \epic\ catalog makes a suitable training set for the main-sequence and sub-giant
branch in \rave.

% Identifying obviously bad stars that have been mis-classified by K2/EPIC
% Requiring stars with narrow posteriors.
% Make sure you mention the lack of abundance information in K2/EPIC

>>>>

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\galah}{\project{\acronym{GALAH}}}
\newcommand{\ges}{\project{Gaia-ESO}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\aspcap}{\project{\acronym{ASPCAP}}}
\newcommand{\lamost}{\project{\acronym{LAMOST}}}
\newcommand{\hipparcos}{\project{Hipparcos}}
\newcommand{\epic}{\project{K2/EPIC}}

@@

\begin{abstract}
The Milky Way may be our greatest laboratory for understanding galaxy formation.
This premise relies on the ability to precisely determine the astrometry and
chemistry for many individual stars, and to use those data to make inferences
about the Galaxy's formation and evolution. On 14 September 2016 the \gaia\ 
mission will release precise astrometry for $\approx2\times10^6$ stars.  
However, a comprehensive catalog of detailed chemical abundances to complement
those data is currently lacking.  The RAdial Velocity Experiment (\rave) survey 
has acquired spectra for 292,036 stars in common with the first \gaia\ data 
release, constituting the largest overlap of any stellar spectroscopic sample.  
Here we perform an independent analysis of \rave\ spectra using an implementation
of \thecannon\ that uses prior probabilities on atomic line formation and stellar
astrophysical parameters.  Our approach yields precise stellar labels at lower 
S/N ratios than most \emph{ab initio} approaches, allowing us to deliver improved
radial velocities, effective temperature $\teff$, surface gravity $\logg$, and 
chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) for 
\Nstars\ stars.  We validate our results internally from repeat visits, and 
externally against existing high-resolution spectroscopic surveys.  The typical
precision in chemical abundances is X.XX~dex.  We provide spectrophotometric 
distances based on our labels, improving upon the parallax distances available
for the fainter targets in the first \gaia\ data release. \stub{A one-line 
science result based on RAVE/Hipparcos overlap}
\end{abstract}

@@

The Milky Way is considered to be the best laboratory for understanding galaxy
formation and evolution.  This premise hinges on the ability to precisely measure 
the astrometry and chemistry for (many) individual stars, and to use those data 
to infer the structure, kinematics, and chemical enrichment of the Galaxy 
\citep[e.g.,][]{Schlaufman_2009,Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,
Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Bovy_2016}.  However, these 
quantities are not known for even 1\% of stars in the Milky Way.  Stellar 
distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,Jofre_2015,
Madler_2016}, proper motions can be plagued by unquantified systematics from 
the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
stellar spectroscopists frequently report significantly different chemical 
abundance patterns \citep{Smiljanic_2014}.  The impact these issues have on our
scientific inferences cannot be understated.  They limit our understanding on a
number of sub-fields in astrophysics, including the properties of exoplanet host
stars, the formation (and destruction) of stars and clusters, as well as the 
structure of the Galactic disk, to name a few.

These formed just some of the scientific drivers for the \gaia\ space telescope.
\gaia\ is primarily an astrometric mission, providing precise positions,
parallaxes and proper motions for more than $10^9$ stars in its final data
release in 2022.  While this is a sample size about three orders of magnitude 
larger than its predecessor \hipparcos, both astrometry and chemistry are 
required to fully characterize the formation and evolution of our Galaxy.  For
this reason \gaia\ will also provide radial velocities, stellar parameters and
chemical abundances for a subset of brighter stars.  While astrometry for bright
and nearby stars will be imminently available, chemical abundances from \gaia\ 
will not be available until later releases.  Until those abundances are available,
astronomers seeking to use chemistry and kinematic measurements are reliant on 
ground-based spectroscopic surveys in order to complement the available astrometry.

The first \project{Gaia} data release will include positions, proper motions, and 
parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} 
catalog \citep{Michalik_2015a,Michalik_2015b}.  After cross-matching all major 
stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2
catalog against the \apogee\ \citep{Zasowski_2013}, \ges\
\citep{Gilmore_2012,Randich_2013}, \galah\ \citep{DeSilva_2015},
\lamost\ \citep{Cui_2012}, and \rave\ \citep{Steinmetz_2006} 
surveys}, we found the RAdial Velocity Experiment (hereafter \rave) survey to 
have the largest overlap with the first \gaia\ data release: 292,036 
stars.  We then used the \gaia\ universe model snapshot \citep{Robin_2012} to 
estimate the precision in parallax and proper motions that will be available in 
the first \gaia\ data release (DR1) for stars in those overlap samples. Comparing
the expected precision to what is currently available, we further found that the
\rave\ survey will benefit most from \gaia\ DR1.  The distances of 63\% of the 
\rave--\gaia\ DR1 overlap sample (182,862 stars) are expected to improve with 
the first \gaia\ data release, and 47\% of stars are likely to have better 
proper motions (137,211 stars).  This motivated us to examine what chemical 
abundance information was available for \rave, and to evaluate whether we could
enable new chemodynamic studies by contributing to this existing set of chemical
abundances.

We briefly describe the \rave\ data in Section~\ref{sec:data}, before explaining
our method in Section~\ref{sec:method}.  In Section~\ref{sec:validation}
we outline a number of validation experiments, including: internal sanity checks,
comparisons with literature samples, and investigations to ensure our results
are consistent with expectations from astrophysics.  We discuss these results in
Section~\ref{sec:discussion}, and in Section~\ref{sec:conclusion} we provide 
concluding remarks as well as a description of how to access our results electronically.

@@

\rave\ is a magnitude-limited stellar spectroscopic survey of the (nearby) Milky Way,
principally designed to measure radial velocities for up to $10^6$ stars.
Observations were conducted on the 1.2~m UK Schmidt telescope at the Australian 
Astronomical Observatory\footnote{Formerly the Anglo-Australian Observatory} from 
2003--2013.  The large 5.7~degree field-of-view and robotic fibre positioner (6dF)
made for very efficient observing.  Spectra for up to 150 targets were simultaneously
acquired, and when observations concluded in April 2013, a total of 574,630 spectra
had been collected of 483,330 unique objects.

The target selection for \rave\ is based on the $I$-band apparent magnitude,
$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge \citep{Wonjo_2016}.  
The $I$ band was selected because it overlaps with the wavelength range that \rave\ 
operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} near-infrared 
triplet, and a number of atomic transitions of light-, $\alpha$-, and Fe-peak elements.
The \ion{Ca}{2} near infrared triplet lines are dominated by pressure broadening, 
which makes them strong and therefore visible even in very low S/N data or 
metal-poor stars.  This spectral region is also very well-studied.  It is one of the
key setups used for the ground-based high-resolution \ges\ survey
\citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
coverage is comparable to the Radial Velocity Spectrometer on board the \gaia\
space telescope \citep{Recio-Blanco_2016}.

The exposure times for \rave\ observations were optimised to obtain radial 
velocities for as many stars as possible.  Detailed chemical abundances were
always an important science goal of the survey, but this was a secondary objective.  
For this reason the distribution of S/N ratios in \rave\ spectra is considerably 
lower than other stellar spectroscopic surveys where chemical abundances are the 
primary motivation.  The \rave\ spectra have an effective resolution 
$\mathcal{R} \approx 7500$ and the distribution of S/N ratios peaks at 
$\approx$50~per~pixel$^{-1}$.  For comparison, the \galah\ survey 
\citep{DeSilva_2015} --- which was specifically constructed for detailed chemical 
abundance analyses --- includes a wavelength range about 2.5 times larger at 
resolution $\mathcal{R} \approx 28,000$, and yet the \galah\ project still 
targets for S/N $\gtrsim100$.

@@

Given the high-quality of the normalization performed by the \rave\ team, we chose
not to re-normalize the spectra.  Our tests demonstrated that the procedure 
outlined in \citet{Kunder_2016} is sufficient for our analysis procedure. Therefore,
there is a limited number of pre-processing steps that we conducted before our
analysis could begin.  Firstly, we calculated inverse variance arrays from the
fractional uncertainties provided, then re-sampled the flux and inverse variance
arrays onto a common rest-wavelength map for all stars.  Depending on the fibre 
used and the stellar radial velocity, the range of rest-frame wavelength values
varied for each star.  Given that fluxes were unavailable in the edge pixels for 
most stars, we excluded pixels outside of the rest wavelength range 
$8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.  This corresponds to about
30~pixels excluded on either side of the common wavelength array, leaving us with
945~pixels per spectrum for science.

@@

Throughout this work we adopt the assumptions and nomenclature first set out in
\citet{Ness_2015}, and extended in \citet{Casey_2016b}.  However, one distinction 
here with \citet{Casey_2016b} is that in this work the \emph{labelled} set and the
\emph{training} set will be the same set.  In \citet{Casey_2016b} two subsets of
the labelled set were defined: the training set and a distinct validation set.
This allowing us to train on the \emph{training set} and then test on the 
\emph{validation set}.  However in this work the labelled set is relatively 
small, about an order of magnitude smaller than what was available in 
\citet{Casey_2016b}, and therefore no validation set exists. We will use all 
labelled set stars in the training step and supplement our validation experiments 
with literature comparisons.  Thus, throughout this work we will use the terms 
\emph{labelled set} and \emph{training set} interchangeably.  Below we describe 
how we constructed the training set, before we outline the model that we adopt.

@@

We sought to construct a training set consisting of stars across the main-sequence,
sub-giant, and the red giant branch.  We required stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \ges\ survey includes 
giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
\apogee\ data release \citep{sloan_dr13} reports labels for stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the main-sequence labels suffered from significant systematic effects.  A flat, 
then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars incorrectly sit above the isochrones in a classical Hertzsprung-Russell diagram)
If we consider lower-resolution studies, there are 2,369 stars that overlap with 
\lamost\ --- of which 2,213 have positive S/N ratios in the $g$-band 
(\texttt{snrg}).  However, the labels are expectedly less precise given the lower 
resolution, no detailed abundance information is available for main-sequence stars
\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
but that sample contains only giant stars}, and similar systematic trends are apparent
in the lower main-sequence.  The \galah\ survey has a good overlap with \rave, but 
results from that sample are not yet precise enough for our purposes.

These constraints forced us to construct a heterogeneous training set.  Given previous
successes in transferring high S/N ratio labels from \apogee\ \citep{Ness_2015,
Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \apogee---\rave\ 
overlap sample for giant star labels in the training set.  Of these, about 900 are 
giants according to \apogee.  From this sample we selected stars to have: metallicity
determinations ([Fe/H] $> -5$); high S/N ratios in both the \apogee\ and \rave\ spectra
($>200$ in \apogee; $>50$ in \rave); and we further required that the \aspcap\ pipeline 
did not report any peculiar flags (\texttt{ASPCAPFLAG = 0}).  These restrictions left 
us with 357 stars along the giant branch, with metallicities ranging from 
$[{\rm Fe/H}] = -1.62$ to 0.32.  Intermediate tests with globular cluster members showed
that the metallicity range of the training set needed to extend below 
$[{\rm Fe/H}] \lesssim -2$ in order for our catalog to be practically useful.
Therefore, as we will describe in more detail below, we supplemented this sample with 
giant stars from the \epic\ catalog \citep{Huber_2016}.

Assembling a training set for the main-sequence and sub-giant branch was more complex.
There are no spectroscopic studies that have a large enough sample size that overlap 
with \rave, and which also cover our required range in stellar parameters 
(e.g., FGKM-type stars).  Moreover, most of the spectroscopic studies we considered 
exhibit a flat lower main-sequence, a systematic consequence of the analysis method 
adopted \citep[e.g., see][for a discussion]{Bensby_2014}.  For these reasons we chose
to make use of the \epic\ catalog \citep{Huber_2016} for the training set labels on
the main-sequence and sub-giant branch.  The \epic\ catalog follows from the successful
\project{Kepler} input catalog \citep{Brown_2011}, and provides probabilistic stellar 
classifications for 138,600 stars in the \project{K2} fields based on the 
astrometric, asteroseismic, photometric, and spectroscopic information available for
every star.  There are 4,611 stars that overlap between \epic\ and \rave.

\epic\ differs from the \project{Kepler} input catalog because \epic\ does not 
benefit from having narrow-band $DDO_{51}$ photometry in order to aid dwarf/giant 
classification.  Despite this limitation the labels in the \epic\ catalog have 
already been shown to be accurate and trustworthy \citep{Huber_2016}.  However, 
when the posteriors are wide (i.e., the quoted confidence intervals are large) 
due to limited information available, it is possible that a star has been 
mis-classified.  This is most prevalent for sub-giants, where \citet{Huber_2016} 
note that $\approx55-70$\% of sub-giants are misclassified as dwarfs.  The 
probability of misclassification is usually quantified in the uncertainties of 
each star; most sub-giants that could be dwarfs have large confidence intervals.  
Therefore, requiring low uncertainties will decrease the total sample size, but 
will remove most misclassifications.  The situation is far more favourable for 
dwarfs and giants.  Only 1-4\% of giant stars are misclassified as dwarfs, and 
about 7\% of dwarfs are misclassified as giants.  To summarise, the \epic\ labels 
with narrow confidence intervals are usually of high fidelity, and any spurious 
misclassifications would be clearly distinguishable if spectra were available.

We sought to have a small overlap between our giant and main-sequence star training
sets.  Most of our giant training set is encapsulated within $0 < \logg < 3.5$, 
however there is a sparse sampling of stars reaching to $\logg \approx 4$.  For
the \epic\ main-sequence/sub-giant star training set we required $\logg > 3.5$,
allowing for $\approx0.5$~dex of overlap between the two training sets.  We further
employed the following quality constraints: the upper and lower confidence intervals 
in $\teff$ must be below 150~K; the upper and lower confidence intervals in $\logg$ 
must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must exceed 
30~per~pixel$^{-1}$; and $\teff \leqslant 6750$.  Unfortunately these strict constraints
removed most metal-poor stars, which we later found to cause the test labels to have
under-predicted abundances for dwarfs of low metallicity.  For this reason we relaxed
(ignored) those quality constraints for stars in the \epic\ catalog with 
$[{\rm Fe/H}] < -1$.  Of the stars we forcibly included with $[{\rm Fe/H}] < -1$,
none were classified as being sub-giants in \epic, therefore they have a low (7\%)
likelihood of misclassification.  The distilled sample of the \rave--\epic\ overlap
catalog contains 583 stars (of 4,611).

% We can also identify misclassified stars by cross-validation.

% Make sure you mention the lack of abundance information in K2/EPIC
